---
title: "CLI Guide"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with httr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE}
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)
knitr::opts_chunk$set(comment = "#>", collapse = TRUE)
```

# BaMORC CLI guide

So you want to using `BaMORC` in shell no matter linux or windows? This document walks through each command line functions built-in the `BaMORC` package.

## Over all CLI design

CLIs vary widely. Before diving into the actual command, it is important to understand the general patter of the CLI commands so that you can construct a complete and coherent command for utilizing `BaMORC` funcationalities.

The key features of any CLI is the structre of the command. A CLI consists of the following paters:

1. Language interpreter,
2. A script for doing something,
3. Arguments that need to pass in the script.

You need to be able to constructure a command with these components in order to perform the desired CLI call.

`BaMORC` supports two type of CLI calls: innate `R CMD BATCH`, which supports all OS platforms, and Linux CLI.

### R CMD BATCH (Windows-friendly)

R language has its own CLI command pattern via the functionality of `R CMD BATCH`. At the very beginning of R history, `R CMD BATCH` was the function handle all the non-interactive calls through the shell. 

   The patters is simple: `R CMD BATCH [arguments-to-pass] [path-of-script.R] $` 
   
Here arguments are in the `-` format, for example `-ppm_range`. More detail refer to [Argument session](## Arguments to pass in the R script)

### Linux CLI

The standard Linux CLI follows the `Docopt` convention. 

   The pattern is: `[path-of-script.R] [command] [arguments-to-pass]`

The command in the Linux CLI is either `assigned`, `unassigned` or `valid_ids`. The arguments are in `--` format, for example `--ppm_range`. More detail refer to [Argument session](## Arguments to pass in the R script) 


## Arguments and examples

### R CMD BATCH
Using the `R CMD BATCH` will allow interactive, aka. after call the R script with argument, the terminal will not wait the script finishing, so no print-out no result will show in terminal window. In order to veiw the results, you need to find the output file saved on the location you provided.
```{r}
library(BaMORC)
```


### Linux CLI